Ionospheric Disturbances in Mexican Territory Produced by Objects Entering the Athmosphere from Space \tnoteref{tnote1}

### 2021-09-02 paper/manuscript.tex
Newest changes in scripts, new scripts and advances in paper
- Slight improvements to existing scripts
- Add new script where filters PRNs into a defined time interval around the moment the meteor is fragmented
- Split sections of paper into separated files
    - Advances in methodology section

\verso{Given-name Surname \textit{etal}}

@@

%% main text
\section{Introduction}
\label{sec1}
%Please use \verb+elsarticle.cls+ for typesetting your paper. Additionally,
%make sure not to remove the package \verb+jasr.sty+ already included in the
%preamble:
%\begin{verbatim} 
%  \usepackage{jasr}
%\end{verbatim}

@@

\section{Metodology}
\subsection{Meteors Database}

We selected a sample of meteors which were observed in mexican territory from the Geostationary Lightning Mapper \citep{GOODMAN:2013}. Orignally this project was designed to detect ligthning activity in earth's athmosphere, but has been proven that also can detect bolides entering the athmosphere. The detection comes from two satellites called GOES-16 and GOES-17 orbiting the earth in geostationary orbits. We used the interactive database available at \url{https://neo-bolide.ndc.nasa.gov/#/} to get the events positions presented in this section, as well we obtained data about the bolid trajectory detected and energy released. The sample was chosen following the next criteria:%selected in such way we chose the most probable objects to be detected by GPS sources in mexican territory and its surroundings:

\begin{itemize}
    \item The objects were detected inside mexican territory and its surroundings.
    \item The objects were detected by both satellites GOES 16 and GOES 17 (stereo)
    \item The detection has been assigned a high confidence ratio.
\end{itemize}

\begin{table*}
  \centering
  \caption{List of meteors passing through Mexico. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{|r|r|r|r|r|r|}
\hline
ID & Date of event & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg)\\
\hline
01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$\\
02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$\\
03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$\\
04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$\\
05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$\\
06\tablefootnote{This object was only detected by GOES-16 satellite} & 2019-10-12 & 11:12:27 & $0.006\pm 0.0000$ & $28.8 \pm 0.000$ & $-111.3 \pm 0.000$\\
07 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$\\
08 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$\\
09 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$\\
10 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$\\
11 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$\\
12 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$\\
13 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$\\
14 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$\\
15 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$\\
16 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$\\
17 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$\\
18 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$\\
19 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
20 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$\\
21 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$\\
22 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$\\
23 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$\\
24 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$\\
25 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$\\
26 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$\\
27 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$\\
28 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$\\
29 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
30 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$\\
31 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$\\
32 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$\\
33 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$\\
34 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$\\
35 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$\\\hline
\end{tabular}
\end{table*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{../meteors_map}
  \caption{Positions of events from table \ref{tab:table-meteors}. The label of each point correspond to the ID (first column) of the referred table.}
  \label{fig:meteors-map}
\end{figure}

\subsection{GPS data}

For the selected sample, we obtained RINEX data from the TlalocNet \citep{Cabral-Cano:2018} and UNAVCO network databases to study potential alterations in the ionospehre due to the presence of the passing meteor at the day the meteor was reported. For each event, we downloaded data from stations that surrounds the place where the event was detected (usually 3 to 5 stations. )The list of the sample meteors is shown in table \ref{tab:table-meteors}. The events are in chronological order. The reported duration, latitude and longitude correspond to the mean between measurements from satellites GOES-16 and GOES 17; in the same way, the uncertainties correspond to the standard deviation. Also their respective positions are available in figure \ref{fig:meteors-map}, where each label correspond to the ID (first column) of table \ref{tab:table-meteors}.

Using the data provided by TlalocNet and UNAVCO, we procceeded to obtain TEC parameters with the GPS\_GOPI software, available at \url{https://seemala.blogspot.com/}. This software takes as input the RINEX data (the navigation file is no strictly neccesary), and the outuput consists in in the vTEC and sTEC measurements for the PRNs of the whole day the event occurred, as well as the averaged TEC as function of time. We obtained vTEC maps for the events in table (\ref{tab:table-meteors}) and their respective next and previous day.

%Final idea: use GOPI software to get the vTEC data from the day of each event and the previous days.

\section{Ionospheric background and vTEC maps}

Ionospheric perturbations also can take place due to space weather and geomagnetic storms. So, in order to discard such events we investigated the space weather in the day each event occured. In figure (name) we present the geomagnetic \textit{Kp} index for some events. We discarded events whose Kp index is equal or grater than 4 in the day of the event or shortly before. Also we present in figure (name) the vTEC perturbation maps for the same events in a three day series, centered in the event date. The estimated meteor trajectory, obtained from the GLM data is presented in black, continous line, while the linear fit to the GOES-16 and GOES-17 data are presented with the red dashed line, and work as boudary errors.

\section{Frecuency Analysis}

\section{Discussion}

%\section{How to include a figure?}
%Change the file name of the figure below with your own 
%figure. And remember to change the figure caption too!
%\begin{verbatim}
%\begin{figure}
%  \centering
%  \includegraphics[scale=0.5]{fig01}
%  \caption{Write the figure caption here.}
%  \label{fig:pendulum}
%\end{figure}
%\end{verbatim}

%\begin{figure}
%\centering
%\includegraphics[scale=0.5]{fig01.pdf}
%\caption{Write the figure caption here.}
%\label{fig:pendulum}
%\end{figure}

@@

\section{Acknowledgments}

The RINEX data used in this paper were obtained from the Trans-boundary, Land and Atmosphere Long-term Observational and Collaborative Network (TLALOCNet, \citet{Cabral-Cano:2018}), operated by the Servicio de Geodesia Satelital (SGS) at the Instituto de Geofísica-Universidad Nacional Autónoma de México (UNAM) in collaboration with UNAVCO Inc.

We are deeply grateful to all personnel from SGS, SSN and UNAVCO for station installation, maintenance, data acquisition, IT support and data curation and distribution for these networks and in particular to the following individuals and institutions, and those whose hard field work and resourcefulness were central to the success of this project: Bill Douglass, Neal Lord and Bill Unger at UW-Madison, Oscar Diaz-Molina and Luis Salazar-Tlaczani at SGS, John Galetzka, Adam Wallace, Shawn Lawrence, Sean Malloy and Chris Walls at UNAVCO, Jesus Pacheco-Martínez at Universidad Autónoma de Aguascalientes, Bertha Marquez-Azúa and personnel at the Universidad de Guadalajara at campus Guadalajara, Mascota and Ameca, Protección Civil de Jalisco, Universidad de Colima at campus Colima and campus El Naranjo and Centro de Geociencias, Centro de Ciencias de la Atmosfera Instituto de Biología Estacion Chamela at UNAM.
TLALOCNet, SSN-TLALOCNet and other GPS related operations from SGS are supported by CONACyT projects 253760, 256012 and 2017-01-5955, UNAM-PAPIIT projects IN104213, IN111509, IN109315-3, IN104818-3, NSF grant 2025104, NASA-ROSES grant NNX12AQ08G and supplemental support from UNAM-Instituto de Geofísica. UNAVCO's initial support for TLALOCNet (some of its stations now part of the GAGE Facility-NOTA) was performed under EAR-1338091 and is currently supported by NSF and NASA under NSF Cooperative Agreement EAR-1724794.

Also we are grateful with the Direcci\'on General de Asuntos Acad\'emicos (DGAPA) of the Universidad Nacional Aut\'onoma de M\'exico and the Programa de Becas Posdoctorales for funding this project.

%% Bibliography
%% Author year style
\bibliographystyle{model5-names}
\biboptions{authoryear}
\bibliography{bibliography}

>>>>

\verso{Tarango-Yong \textit{et. al}}

@@

%% main text
\section{Introduction}
\label{sec1}
\input{introduction}
%Please use \verb+elsarticle.cls+ for typesetting your paper. Additionally,
%make sure not to remove the package \verb+jasr.sty+ already included in the
%preamble:
%\begin{verbatim} 
%  \usepackage{jasr}
%\end{verbatim}

@@

\section{Methodology}
\label{sec:methodology}
\input{methodology}

\section{Bolides physical parameters}
\label{sec:bolides}
\input{meteor-properties}
\section{Ionospheric background and vTEC maps}
\label{sec:vTEC maps}

\input{vTEC-maps}

%\section{Frecuency Analysis}

\section{Discussion}
\label{sec:discission}

\input{discussion}
%\section{How to include a figure?}
%Change the file name of the figure below with your own 
%figure. And remember to change the figure caption too!
%\begin{verbatim}
%\begin{figure}
%  \centering
%  \includegraphics[scale=0.5]{fig01}
%  \caption{Write the figure caption here.}
%  \label{fig:pendulum}
%\end{figure}
%\end{verbatim}
%\begin{figure}
%\centering
%\includegraphics[scale=0.5]{fig01.pdf}
%\caption{Write the figure caption here.}
%\label{fig:pendulum}
%\end{figure}

@@

\section{Acknowledgments}
\label{sec:acknowledgments}
\input{Acknowledgments}
%% Bibliography
%% Author year style
\bibliographystyle{model5-names}
\biboptions{authoryear}
\bibliography{bibliography}
